---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Loading and preprocessing the data

```{r}
activity = read.csv("activity.csv")

data <- na.omit(activity)
```


## What is mean total number of steps taken per day?

```{r}
stepsByDate <- aggregate(steps ~ date, data = activity, FUN = sum)
barplot(stepsByDate$steps, names.arg = stepsByDate$date, xlab = "date", ylab = "steps")

mean <- format(mean(stepsByDate$steps), nsmall = 2)
median <- format(median(stepsByDate$steps), nsmall = 2)
```

The mean total number of steps taken per day: <b> `r mean`</b>

The median total number of steps taken per day: <b>`r median`</b>

## What is the average daily activity pattern?
```{r}
stepsInterval <- aggregate(steps ~ interval, data = activity, FUN = "mean")
plot(stepsInterval, type = "l")

max <- stepsInterval$interval[which.max(stepsInterval$steps)]
```

Interval that contains the maximum number of steps: <b>`r max`</b>

## Imputing missing values

```{r}
nas <- sum(is.na(activity))
```


Total number of missing values: <b>`r nas`</b>

```{r}
activity <- merge(activity, stepsInterval, by = "interval", suffixes = c("", 
    ".y"))
nas <- is.na(activity$steps)
activity$steps[nas] <- activity$steps.y[nas]
activity <- activity[, c(1:3)]
```

```{r}
stepsBydate <- aggregate(steps ~ date, data = activity, FUN = sum)
barplot(stepsBydate$steps, names.arg = stepsBydate$date, xlab = "date", ylab = "steps")

newMean <- format(mean(stepsBydate$steps), nsmall = 2)
newMedian <- format(median(stepsBydate$steps), nsmall = 2)
```

New mean: <b>`r newMean` </b>

New median: <b>`r newMedian` </b>


## Are there differences in activity patterns between weekdays and weekends?

```{r}
daytype <- function(date) 
{
    if (weekdays(as.Date(date)) %in% c("Saturday", "Sunday")) 
    {
        "weekend"
    }
    else 
    {
        "weekday"
    }
}

activity$daytype <- as.factor(sapply(activity$date, daytype))

par(mfrow = c(2, 1))

for (type in c("weekend", "weekday")) 
{
    steps.type <- aggregate(steps ~ interval, data = activity, subset = activity$daytype == 
        type, FUN = "mean")
    plot(steps.type, type = "l", main = type)
}

```

